Explorative Data Analysis Climate Perceptions in Countries

### 2022-11-28 Code/Drafts/Draft_2_GHG_Emissions_&_Perceptions_of_Climate_Change.Rmd
moving files making draft folder

I might use active voice "I will analyze data" vs. "data will be analyzed"

@@

In the emissions section Denmark's and the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U. S. had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]

Go ahead and use "I" rather than "We" because this is  your work.

@@

Good overall, but also include links to each dataset.

@@

Now tell R the year column are dates so we can better perform statistical analysis.

Below works fine, but just so you are aware, it is base R and not lubridate.

```{r}
emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
#format tells R the dates are just a four digit year, a two digit year is %y, origin informs R where it starts.
# must do the - in the %d- because the dates are entered with a -. 
```

@@

Already it is apparent that Denmark produces less GHG emissions per capita each year than the United States. Both countries have been decreasing their emissions since 1990, but  Denmark seems to have a  greater rate of decrease in emissions, especially since 2006, than the United States.

Running a _______ statistical test will let us compare the rate of decrease in emissions of both these countries.

I talked to Robin Lock about this. We're going to do a paired t-test where each year has a pair of values (US, Denmark) and the values will be the RATE of change for that year, rather than the GHG emissions themselves.

Here's some info: https://www.statology.org/paired-samples-t-test-r/

We can also compare the moving averages and compare the rates, even forecasting to future dates (like what will emissions be in 2050 or 2100 based on the current time series). This comparision will be visual for the moving averages and just thoughtful for comparing the forecast numbers and will not involve a specific statistical test.

@@

First determine the minimum date
```{r}
min_date=min(emissions$year)
max_date=max(emissions$year)
```
I'm not sure you'll need the above code chunk

@@

If you wish to plot time series using ggplot, you can use autoplot(timeseriesname) and save it as an object. The result will be a ggplot object that you can manipulate. Example:

```{r}
demo_plot<-autoplot(emissions_US_ts)
demo_plot+
  labs(x = "Year", y = "Greenhouse gas emissions")+
  theme_bw()
```

And if you want to plot two time series on the same graph using ggplot, you use cbind to bind your two time series objects into the same new data frame and then autoplot that data frame:

```{r}
both<-cbind(emissions_Denmark_ts, emissions_US_ts)
autoplot(both)
```

@@

plot(emissions_US_ts)
lines(Ma_emissions_US, col="blue")
```
Plotting both with autoplot/ggplot:

```{r}
USA<-cbind(emissions_US_ts, Ma_emissions_US)
autoplot(USA)
```

@@

***PUT RUNNING THE TIME SERIES HERE***
This will really be the paired t-test

@@

I talked with Prof Shuckers about these data. He thinks we should do the back-calculating to change percents to numbers and then do a chi-square to compare answers between Danes and Americans.  You will need the data on number of respondants for each question to do that.

@@

If you want to test these questions with a statistical test, we will need a combination of chi-square tests. Let's talk.

@@

Here's an article that talks more about data viz for likert type data. It has good arguments for why to avoid pie charts, and has a building series of charts ending with the author's (and my) favorite. Just worth considering:

https://nightingaledvs.com/visualizing-likert-scale-data-same-data-displayed-seven-different-ways/

@@

Chi-square test after converting percents to numbers by multiplying the percents times the total number of repsondants to that question per country.

@@

#### Lets do a statistical test to verify. 
Chi-square test. And think about a new data vis.

@@

parsing the data
```{r}
threat<- survey %>% filter(survey$question=="Ques__threat_20_years") 
```

@@

What about challenges in terms of figuring out how to reshape/reformat your data? Anything worth mentioning there?

This is very thorough and a solid first draft.

>>>>

@@

In the emissions section Denmark's the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U. S. had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]

@@

@@

Now tell R the year column are dates so we can better perform statistical analysis. 
```{r}
emissions$year<-as.Date(emissions$year, format="%d-%m-%Y")
#format tells R the dates are just a four digit year, a two digit year is %y, origin informs R where it starts.
# must do the - in the %d- because the dates are entered with a -. 
```

@@

Already it is apparent that Denmark produces less GHG emissions per capita each year than the United States. Both countries have been decreasing their emissions since 1990, but  Denmark seems to have a  greater rate of decrease in emissions, especially since 2006, than the United States. Running a _______ statistical test will let us compare the rate of decrease in emissions of both these countries.

@@

First determine the minimum date
```{r}
min_date=min(emissions$year)
max_date=max(emissions$year)
```

@@

@@

plot(emissions_US_ts)
lines(Ma_emissions_US, col="blue")
```

@@

***PUT RUNNING THE TIME SERIES HERE***

@@

@@

@@

@@

@@

#### Lets do a statistical test to verify. 
Again we will do a general linear model.

@@

parcing the data
```{r}
threat<- survey %>% filter(survey$question=="Ques__threat_20_years") 
```

@@

### 2022-11-28 Code/Drafts/Draft_Exploratory_Analysis_CAP_DAP.Rmd
moving files making draft folder

## Yale Survey Data
Bring in the data
```{r}
yale<- read.csv(here("Data for the Code", "Yale_International Perspective of Climate Change.csv"))
```

>>>>

## Yale Survey Data
Bring in the data
```{r}
yale<- read.csv(here("Data for the Code", "Raw Data Sets", "Yale_International Perspective of Climate Change.csv"))
```
